Pre-processing high frequency animal tracking data

### 2020-10-07 02_filtering_data.Rmd
some code edits
add some text

# Filtering data

@@

@@

# to simulate movement
library(smoove)

# to generate a landscape
library(raster)
library(NLMR)

# strings
library(glue)

@@

## Introducing errors {-}

@@

```{r add_outlier}
# make a dopy
data_copy <- copy(data)
# add a reflection
data_copy[500:800, `:=`(x = x + 0.25,
                        y = y + 0.25)]
# add normal error
data_copy[, `:=`(x = do_add_error(x, std_dev = 0.01),
                 y = do_add_error(y, std_dev = 0.005))]
# add 100 outliers
data_copy <- do_add_outliers(data_copy, p_data = 0.005, std_dev = 0.1)
```

```{r}
# define red and blue
pal <- RColorBrewer::brewer.pal(4, "Set1")
```

## Show data with errors

```{r}
# make figure
figure_raw <-
  ggplot()+
  geom_point(data = data_copy,
            aes(x, y),
            col = "grey",
            alpha = 1,
            size = 0.2)+
  geom_path(data = data,
            aes(x, y),
            col = "grey20",
            alpha = 1)+
  ggthemes::theme_map()+
  coord_equal()+
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        legend.position = "none")+
    theme(plot.background = element_rect(fill = NA))+
  labs(colour = NULL)
```

@@

First we remove gross positioning errors using a bounding box filter, i.e., knowing or assuming the real bounds of the movement track (in this case, 0 -- 1), we remove all positions outside those bounds.

@@

>>>>

# Reducing Large-Scale Positioning Error by Filtering

@@

Here we load some useful libraries, and the helper functions.

@@

@@

## Introducing Errors to the Data

Here we introduce three kinds of errors to the data:

1. Small-scale normally distributed errors at each position;

2. Large-scale error at a random 0.5% of positions;

3. A large-scale displacement of a sequence of 300 positions.

While the data are 10,000 positions at 1-second interval, we shall use only 5,000 of these.

@@

```{r add_outlier}
# make a copy
data_copy <- copy(data)

# add a prolonged spike or reflection to 300 positions
data_copy[500:800, `:=`(x = x + 0.25,
                        y = y + 0.25)]

# add normal error
data_copy[, `:=`(x = do_add_error(x, std_dev = 0.01),
                 y = do_add_error(y, std_dev = 0.005))]

# add 100 outliers
data_copy <- do_add_outliers(data_copy, p_data = 0.005, std_dev = 0.1)
```

Save the data to which errors have been added.

```{r}
fwrite(data_copy, file = "data/data_errors.csv")
```

Define a palette with 4 colours for convenience.

```{r}
# define a four colour palette
pal <- RColorBrewer::brewer.pal(4, "Set1")
```

## Make Figure of Simulated Data

We make a figure of the canonical data (grey line) along with the artificially added error (grey points).

```{r}
# make figure of canonical data with added errors
figure_raw <-
  ggplot()+
  geom_point(data = data_copy,
            aes(x, y),
            col = "grey",
            alpha = 1,
            size = 0.2)+
  geom_path(data = data,
            aes(x, y),
            col = "grey20",
            alpha = 1)+
  ggthemes::theme_map()+
  coord_equal()+
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        legend.position = "none")+
    theme(plot.background = element_rect(fill = NA))+
  labs(colour = NULL)
```

@@

First we remove gross positioning errors using a bounding box filter using the function `atl_filter_bounds`. In this example, we show filtering only on the Y coordinate.
`atl_filter_bounds` takes coordinate ranges as two-element vectors of the lower and higher bound.
It is possible to pass one the bounds as `NA`, in which case, only the other bound is used for filtering, i.e., `y_range = c(NA, 1)` is equivalent to selecting all Y coordinates < 1.

`atl_filter_bounds` was initially designed to remove positions _inside_ a specific range, hence the argument `remove_inside`.
The default value of the argument is `FALSE`, and `atl_filter_bounds` is thus a bounding box filter.

@@

`atl_filter_bounds` is not vectorised, and if there are two or more bounds per coordinate (for instance, $x_1 \ldots x_2$, and $x_3_ \ldots x_4$), they must be passed in two different function calls.
